Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-25 NEUROC/Response-MajRev/response_letter.tex

     In the end of Section 4.3, we added a new paragraph ``Memory Usage of Inference'' to compare the memory usage of inference and training.
     %
     Figure~\ref{fig:compare_memory_expasion_ratio} in the response (Figure X in the revised manuscript) compares the memory expansion ratios of typical GNNs during training and inference.
     %
     Since no intermediate results had to be cached during inference of GGNN, GAT, and GaAN, the memory expansion ratios of inference were much less than training for the three GNNs.
     %
     The MERs of inference were 45\% to 83\% (GGNN), 52\% to 61\% (GAT), and 37\% to 69\% (GaAN) of training.
     %
     However, the MERs were still high, disallowing inferencing with big graphs.
     %
     To handle big graphs, sample-based inference was necessary.

     \begin{figure}[H]
        \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_memory_expansion_ratio.pdf}}
        \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_memory_expansion_ratio.pdf}}
        \caption{Memory expansion ratios of typical GNNs.}
        \label{fig:compare_memory_expasion_ratio}
     \end{figure}

     \paragraph{Section 4.4 ``Effects of Sampling Techniques on Performance''}
     %
     In Section 4.4, we added a new subsection  Section 4.4.4 ``Performance Bottlenecks in Inference'' to discuss the performance bottlenecks in sample-based inference.

     Sample-based inference was an effective technique to handle big graphs.
     %
     Taking the node classification task as an example, to predict labels for a given set of vertices $\mathcal{V}_{predict}$, the inference sampler sampled a subgraph containing complete $L$-hop neighborhoods of all vertices in $\mathcal{V}_{predict}$, where $L$ was the number of GNN layers.
     %
     Since the real-world graphs often had small-world property, the size of the sampled subgraph increased quickly as the number of vertices in $\mathcal{V}_{predict}$ (i.e., batch size) increased.
     %
     Figure~\ref{fig:sampled_graph_size} (Figure X in the revised manuscript) shows how the relative batch size ($\frac{|\mathcal{V}_{predict}|}{|\mathcal{V}|}$) affected the average degree and the number of edges of the sampled subgraphs during inference.
     %
     When the relative batch size was 10\%, the number of edges in the sampled subgraphs were close to the number of edges in the whole graph in most datasets.
     %
     In order to limit the memory usage during inference, the batch size used should be very small.

     \begin{figure}[H]
        \centering
        \subfloat[Average degree]{    \includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_sampling_graph_info_inference_sampler_avg_degrees.pdf}}
        \subfloat[Number of edges]{    \includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_sampling_graph_info_inference_sampler_edges.pdf}}

        \caption{Characteristics of the sampled subgraphs produced by the inference sampler under different relative batch sizes.}
        \label{fig:sampled_graph_size}
     \end{figure}

     However, the overheads brought by sampling and data transfering from CPU to GPU became obvious when the batch size was small.
     %
     Figure~\ref{fig:time_breakdown_of_inference_sampler} (Figure X in the revised manuscript) shows the time breakdowns of sample-based inference on different datasets.
     %
     On the \texttt{amp} and \texttt{cam} datasets, the sampling time even accounted for near half of the total inference time.
     %
     The results indicated that the current implementation of inference sampler in PyG was also inefficient.
     %
     Improving its efficiency could significantly reduce the inference time.

### 2020-12-25 NEUROC/Response-MajRev/response_letter.tex

% !TEX program=pdflatex
% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{times}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{multirow}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{float}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{framed}
\usepackage[colorlinks=true]{hyperref}
% Vector
\newcommand{\MyVec}[1]{\boldsymbol{#1}}
% Matrix
\newcommand{\MyMat}[1]{\boldsymbol{#1}}
% Vectors with hat
\newcommand{\hvec}[1]{\hat{\boldsymbol{h}}_{#1}}
% Model parameters
\newcommand{\Param}[1]{\textcolor{blue}{#1}}

\usepackage{blindtext}
\usepackage[most]{tcolorbox} 
\definecolor{block-gray}{gray}{0.95}

\newtcolorbox{zitat}[2][]{%
    colback=block-gray,
    grow to right by=-0mm,
    grow to left by=-0mm, 
    boxrule=0pt,
    boxsep=0pt,
    breakable,
    enhanced jigsaw,
    borderline west={4pt}{0pt}{gray},
    title={#2\par},
    colbacktitle={block-gray},
    coltitle={black},
    fonttitle={\bfseries},
    attach title to upper={},
    #1,
}

@@

    \begin{myquote}{Section 4.1 ``Effects of Hyper-parameters on Performance''}
    %
    \emph{Effects on Accuracy}.
    %
    The relationship between hyper-parameters of GNNs (i.e., model complexity) and accuracy is very complicated, and there is no clear theoretical analysis yet.
    %
    Generally speaking, higher model complexity can bring powerful representation capability and may bring higher accuracy, but it also increases the risk of overfitting.
    %
    To evaluate the effects of hyper-parameters on accuracy, we measured the accuracy of the four typical GNNs under different hyper-parameters.
    %
    \figurename~\ref{fig:exp_hyperparameter_accuracy} shows the experimental results.

    For GCN, the accuracy was sensitive to the dimension of hidden vectors $dim(\MyVec{h}^1_x)$.
    %
    As $dim(\MyVec{h}^1_x)$ increased, the accuracy first increased quickly and then stabilized when $dim(\MyVec{h}_x^1) \geq 8$.
    %
    For GGNN, its accuracy curves showed similar trends as GCN.
    %
    However, GGNN was more sensitive to $dim(\MyVec{h}_x^1)$ than GCN.
    %
    Its accuracy even decreased when $dim(\MyVec{h}_x^1) \geq 1024$.
    %
    Since GGNN had high model complexity (with 13 weight matrices/vectors to train), GGNN might occur overfitting in those cases.
    %
    For GAT, its accuracy was more sensitive to the dimension of each head $d_{head}$ than the number of heads.
    %
    For GaAN, only $dim(\MyVec{h}_x^1)$ showed obvious impacts on accuracy.
    %
    The experimental results indicated the accuracy of the GNNs was often low when $dim(\MyVec{h}_x^1)$ (for GCN/GGNN/GaAN) or $d_{head}$ (for GAT) was small.
    %
    In those cases, the low model complexity limited the expressive power of the GNNs.
    %
    As $dim(\MyVec{h}_x^1)$ or $d_{head}$ increased to a certain threshold, the GNNs had sufficient learning ability to achieve stable accuracy.

    \begin{figure}[H]
        \centering
        \subfloat[GCN]{\includegraphics[width=0.4\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gcn.pdf}}
        \subfloat[GGNN]{\includegraphics[width=0.4\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_ggnn.pdf}}\\
        \subfloat[GAT]{\includegraphics[width=0.8\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gat.pdf}}\\
        \subfloat[GaAN]{\includegraphics[width=1.0\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gaan.pdf}}
        \caption{Effects of hyper-parameters on accuracy of the typical GNNs.}
        \label{fig:exp_hyperparameter_accuracy}
    \end{figure}

@@

    %\begin{figure}[H]
        %\centering
        %\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gat_small_info.pdf}
        %\caption{Effects of the number of heads on GAT ($d_{head}$=2,4)}
        %\label{fig:exp_hyperparameter_accuracy_gat_small_info}
    %\end{figure}

    \figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} compares the best accuracy that each GNN achieved on different datasets.
    %
    The best accuracy that the four GNNs achieved was very close.
    %
    It was also close to the accuracy reported in their original references [1][4][6][7].
    %
    There was no clear winner in terms of accuracy.
    %
    The relative accuracy between GNNs varied greatly with different datasets.
    %
    GaAN achieved the highest accuracy in three out of five datasets.
    %
    GCN also achieved the highest or second highest accuracy in three out of five datasets, though its model was simplest.
    %
    It indicated that simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

    \begin{figure}[H]
        \centering
        \includegraphics[width=3.5in]{../figs/experiments/exp_hyperparameter_on_accuracy_alg_contrast.pdf}
        \caption{Best accuracy that each GNN achieved on different datasets.}
        \label{fig:exp_hyperparameter_on_accuracy_alg_contrast}
    \end{figure}

@@

    \end{myquote}
\end{reply}

\begin{point}
    Sampling method is definitely going to reduce model complexity, since all models complexity depend on graph node number N, while performance is going to be compromised as well. I would like to see authors resolve the concern of significant accuracy drop after applying aggressive sampling of subgraphs.
\end{point}

\begin{reply}

    Thank you for your insightful comment and valuable suggestion.
    %
    The sampling methods indeed affect the test accuracy of GNN models \cite{}.% 这里加几个sample方法的文献
    %
    Without sampling, the GNN models are trained in a \emph{batch} gradient descent manner.
    %
    With sampling, the GNN models are trained with \emph{mini-bach} stochastic gradient descent.
    %
    The structure of the GNN remains unchanged between batches.
    %
    Each GNN layer always consists of $|\mathcal{V}|$ graph neurons, where $|\mathcal{V}|$ is the number of vertices in the graph.
    %
    But the numbers of \emph{activated} graph neurons are different for different batches.
    %
    In each batch, only the neurons corresponding to the vertices appearing in the sampled subgraph will be activated.
    %
    The model parameters were updated according to the gradients calculated on the sampled subgraph in each batch.
    %
    Thus, the achieved model accuracy may be different from the full-graph training.

    To discuss how the sampling methods affect model accuracy, we have added a new subsection Section 4.2.2 ``Effects on Accuracy'' in Section 4.4 ``Effects of Sampling Techniques'' in the revised manuscript.
    %
    We quote the added content below.

    \begin{myquote}{Section 4.4.2 ``Effects on Accuracy''}

    Since the size of the sampled subgraph was much smaller than the original graph, the model accuracy got by sampling techniques might be different from the full-batch training.
    %
    To find out how the batch size affected the test accuracy, we trained GNNs with different batch sizes.
    %
    \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} (in the reply) and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} (in the reply) compare the test accuracy achieved by sampling techniques with the test accuracy of the full-batch training.
    %
    For each dataset and each GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section 4.1.

    \begin{figure}[H]
        \centering
        \subfloat[\texttt{amp}]{\includegraphics[width=2in]{../figs/experiments/exp_graphsage_sampling_accuracy_on_amp.pdf}}
        \subfloat[\texttt{amc}]{\includegraphics[width=2in]{../figs/experiments/exp_graphsage_sampling_accuracy_on_amc.pdf}}
        \subfloat[\texttt{fli}]{\includegraphics[width=2in]{../figs/experiments/exp_graphsage_sampling_accuracy_on_fli.pdf}}
        \caption{Test accuracy under different batch sizes of the neighbor sampler. FULL means that the full graph participated in the training.}
        \label{fig:exp_sampling_relative_batch_size_accuracy_graphsage}
    \end{figure}

    \begin{figure}[H]
        \centering
        \subfloat[\texttt{amp}\label{fig:exp_cluster_sampling_accuracy_on_amp}]{\includegraphics[width=2in]{../figs/experiments/exp_cluster_sampling_accuracy_on_amp.pdf}}
        %
        \subfloat[\texttt{amc}\label{fig:exp_cluster_sampling_accuracy_on_amc}]{\includegraphics[width=2in]{../figs/experiments/exp_cluster_sampling_accuracy_on_amc.pdf}}
        %
        \subfloat[\texttt{fli}\label{fig:exp_cluster_sampling_accuracy_on_fli}]{\includegraphics[width=2in]{../figs/experiments/exp_cluster_sampling_accuracy_on_fli.pdf}}
        \caption{Test accuracy under different batch sizes of the cluster sampler. FULL means that the full graph participated in the training.}
        \label{fig:exp_sampling_relative_batch_size_accuracy_cluster}
    \end{figure}

    The experimental results confirmed the effectiveness of the sampling methods in terms of accuracy.
    %
    When the relative batch size was greater than or equal to 3\%, the test accuracy of the GNNs trained with sampling was close to the accuracy obtained by full-batch training.
    %
    In most cases, the accuracy achieved by the sampling techniques was slightly lower than the full-batch training.
    %
    However, there were some exceptions (like GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} and \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc} in the reply) that the accuracy achieved by sampling was even higher.

    The relationships between batch size and test accuracy were complex.
    %
    A larger batch size did not always bring higher accuracy.
    %
    For example, the accuracy of GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} (in the reply) and GGNN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_fli} (in the reply) decreased as the batch size increased.
    %
    A smaller batch size sometimes could achieve higher accuracy.
    %
    For example, GAT achieved a higher accuracy with 1\% relative batch size than the full-batch training in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc} (in the reply).
    %
    Given a sampling method, we found that the optimal batch size highly depended on the dataset and the GNN algorithm.
    %
    Our observations were similar to [31].
    %
    How to automatically select a proper batch size is a topic worth further studying.

    Among the two sampling methods, the performance of the cluster sampler was stabler than the neighbor sampler.
    %
    With the cluster sampler, the test accuracy of different GNNs was very close to the accuracy of full-batch training.
    %
    With the neighbor sampler, the test accuracy of different GNNs showed more obvious differences.
    \end{myquote}

@@

\end{reply}

\begin{point}
    Hope authors supplement the effect of sampling and GNNs on accuracy while comparing different complexity of model and sampling methods.
\end{point}

\begin{reply}

    We are grateful for your insightful suggestion.
    %
    As suggested, we additionally evaluated the effects of model complexity on accuracy in Section 4.1 ``Effects of Hyper-parameters on Performance'' in the revised manuscript.
    %
    We added experimental results on how the hyper-parameters of GNNs affected the accuracy.
    %
    Since different GNNs had different model complexity, we further compared the test accuracy of different GNNs on the same dataset.
    %
    More details are available in the reply to Reviewer Point P2.4.

    We also additionally evaluated the effects of sampling techniques on accuracy in Section 4.4 ``Effects of Sampling Techniques'' in the revised manuscript.
    %
    We presented the experimental results in a new subsection Section 4.4.2 ``Effects on Accuracy''.
    %
    More details are available in the reply to Reviewer Point P2.5.
\end{reply}

### 2020-12-25 NEUROC/Response-MajRev/response_letter.tex

    \begin{figure}[H]
        \centering
        \subfloat[GCN]{\includegraphics[width=0.4\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gcn.pdf}}
        \subfloat[GGNN]{\includegraphics[width=0.4\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_ggnn.pdf}}\\
        \subfloat[GAT]{\includegraphics[width=0.8\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gat.pdf}}\\
        \subfloat[GaAN]{\includegraphics[width=1.0\columnwidth]{../figs/experiments/exp_hyperparameter_on_accuracy_gaan.pdf}}
        \caption{Effects of hyper-parameters on accuracy of the typical GNNs.}
        \label{fig:exp_hyperparameter_accuracy}
    \end{figure}

    %\begin{figure}[H]
        %\centering
        %\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gat_small_info.pdf}
        %\caption{Effects of the number of heads on GAT ($d_{head}$=2,4)}
        %\label{fig:exp_hyperparameter_accuracy_gat_small_info}
    %\end{figure}

    \figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} compares the best accuracy that each GNN achieved on different datasets.
    %
    The best accuracy that the four GNNs achieved was very close.
    %
    It was also close to the accuracy reported in their original references [34, 31].
    %
    There was no clear winner in terms of accuracy.
    %
    The relative accuracy between GNNs varied greatly with different datasets.
    %
    GaAN achieved the highest accuracy in three out of five datasets.
    %
    GCN also achieved the highest or second highest accuracy in three out of five datasets, though its model was simplest.
    %
    It indicated that simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

    \end{myquote}
\end{reply}

\begin{point}
    Sampling method is definitely going to reduce model complexity, since all models complexity depend on graph node number N, while performance is going to be compromised as well. I would like to see authors resolve the concern of significant accuracy drop after applying aggressive sampling of subgraphs.
\end{point}

\begin{reply}

    Thank you for your insightful comment and valuable suggestion.
    %
    The sampling methods indeed affect the test accuracy of GNN models [5, 28, 26, 31].% 这里加几个sample方法的文献
    %
    Without sampling, the GNN models are trained in a \emph{batch} gradient descent manner.
    %
    With sampling, the GNN models are trained with \emph{mini-bach} stochastic gradient descent.
    %
    The structure of the GNN remains unchanged between batches.
    %
    Each GNN layer always consists of $|\mathcal{V}|$ graph neurons, where $|\mathcal{V}|$ is the number of vertices in the graph.
    %
    But the numbers of \emph{activated} graph neurons are different for different batches.
    %
    In each batch, only the neurons corresponding to the vertices appearing in the sampled subgraph will be activated.
    %
    The model parameters were updated according to the gradients calculated on the sampled subgraph in each batch.
    %
    Thus, the achieved model accuracy may be different from the full-graph training.

### 2020-12-25 NEUROC/sec4-experiments.tex
更新R1P6
更新了论文引用的问题

\figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} compares the best accuracy that each GNN achieved on different datasets.
%
The best accuracy that the four GNNs achieved was very close.
%
It was also close to the accuracy reported in their original references \cite{kipf2017_gcn,li2015_ggnn,huang2018_gat,zhang2018_gaan}.
%
There was no clear winner in terms of accuracy.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
GaAN achieved the highest accuracy in three out of five datasets.
%
GCN also achieved the highest or second highest accuracy in three out of five datasets, though its model was simplest.
%
It indicated that simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

\paragraph{GGNN}
%
The top basic operator was \texttt{mm} used in the vertex updating function $\gamma$.
%
Due to its high time complexity, the proportion of \texttt{mm} was much higher than the other operators.
%
The \texttt{thnn\_fused\_gru\_cell} operator was used in the backward phase of $\gamma$.
%
The other three operators were used in the edge calculation stage.

>>>>

\figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} compares the best accuracy that each GNN achieved on different datasets.
%
The best accuracy that the four GNNs achieved was very close.
%
It was also close to the accuracy reported in their original references \cite{shchur2018_pitfall_of_gnn, zeng2020_graphsaint}.
%
There was no clear winner in terms of accuracy.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
GaAN achieved the highest accuracy in three out of five datasets.
%
GCN also achieved the highest or second highest accuracy in three out of five datasets, though its model was simplest.
%
It indicated that simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

\paragraph{GGNN}
%
The top basic operator was \texttt{mm} used in the vertex updating function $\gamma$.
%
Due to its high time complexity, the proportion of \texttt{mm} was much higher than the other operators.
%
The \texttt{thnn\_fused\_gru\_cell} operator was used in both the forward and backward phases of $\gamma$.
%
The other three operators were used in the edge calculation stage.
